---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Loading and preprocessing the data

  Read the data and convert the date column to a date type and the inverval column
to a factor type.

```{r, echo=TRUE}
activity <- read.csv("activity.csv", nrows = 17568)
activity$date <- as.POSIXct(activity$date)
activity$interval <- as.factor(activity$interval)
```

## What is mean total number of steps taken per day?

Calculate the total steps by day using tapply function. Then generate an histogram of the result.

```{r, echo=TRUE}
total_steps <- tapply(activity$steps,activity$date, sum, na.rm = TRUE)
hist(total_steps)
```

Calculate the mean and the median steps by day and print the results.
```{r, echo=TRUE}
mean_steps <- tapply(activity$steps,activity$date, mean, na.rm = TRUE)
median_steps <- tapply(activity$steps,activity$date, median, na.rm = TRUE)

mean_steps
median_steps
```


## What is the average daily activity pattern?

Calculate the average number of steps taken, averaged across all days and then plot the average by 5-minute Intervals.

```{r, echo=TRUE}
mean_interval_steps <- tapply(activity$steps,activity$interval, mean, na.rm = TRUE)
plot(mean_interval_steps, type = "l", xlab = "5-minute interval", ylab = "Average Number of Steps Taken Across All Days", main = "Average Number of Steps Taken by 5-minute Intervals")
```

The 5-minute interval, on average across all the days in the dataset, that contains the maximum number of steps is above.

```{r, echo=TRUE}
mean_interval_steps[mean_interval_steps == max(mean_interval_steps)]
```

## Imputing missing values

Calculate the total amount of NA's in the data and print it.
```{r, echo=TRUE}
totalNA <- sum(is.na(activity))
totalNA
```

Use the the mean of the each 5-minute interval for filling in all of the missing values in the dataset. Put this result in a new variable called activityFill.


```{r, echo=TRUE}
NA_activity <- activity[is.na(activity),]
for (n in 1:nrow(NA_activity)) {
    NA_activity[n,1] <- as.numeric(mean_interval_steps[NA_activity[n,3]])
}

activityFill <- activity
activityFill[is.na(activityFill),] <- NA_activity
```

Calculate the total steps by day using tapply function in the filled data. Then generate an histogram of the result.

```{r, echo=TRUE}
total_stepsFill <- tapply(activityFill$steps,activityFill$date, sum, na.rm = TRUE)
hist(total_stepsFill)
```

Calculate the mean and the median steps by day in the filled data and print the results.

```{r, echo=TRUE}
mean_stepsFill <- tapply(activityFill$steps,activityFill$date, mean, na.rm = TRUE)
median_stepsFill <- tapply(activityFill$steps,activityFill$date, median, na.rm = TRUE)

mean_stepsFill
median_stepsFill
```

Comparing the results of the filled data with the original data, we can see that the values differ from the estimates from the first part of the assignment.
The estimates of the total daily number of steps increased.

## Are there differences in activity patterns between weekdays and weekends?
